### 2014-05-13 Paper.tex
updating Paper.tex

\documentclass[12pt]{article}
\usepackage{amsmath,amsthm,amssymb}
\pagestyle{myheadings}
\markright{Joseph McDonald}

\textwidth=7in
\textheight= 9in

\topmargin=-0.5in
\oddsidemargin=-0.25in

\newcommand{\xbar}{\bar{x}}
\newcommand{\V}[1]{\boldsymbol{#1}}
\newcommand{\E}[0]{\mathbb{E}}

\begin{document}
\title{Online Learning Your arXiv Preferences}
\author{Steve Delong and Joey McDonald}
\maketitle
\begin{abstract}
Our goal was to construct a tool that feeds a user a selection of abstracts of papers from the online database, arXiv.org, recieves a label indicating whether the user is interested in that article or not, and progressively learns the users preferences through an online learning algorithm. It uses this information to present articles that it believes the user will like. We created and tested this using a few sample users' preferences including each of ours. Beyond constructing the tool we found some theoretical justification for bounding the number of updates in expectation while only presenting a fraction of samples that are predicted to be uninteresting.
\end{abstract}

\section{Introduction}

The online database arXiv receives several thousand new submissions a month, and sifting through those articles even in just a single subfield to find one of particular interest can be a daunting task. Since arXiv conveniently stores metadata of its submissions in a fairly accessible XML format, a program that compiles and uses this metadata to find papers matching particular research interests could be a great tool for the scientists intending to peruse new research. Essentially this would serve as a ``Pandora," the music recommendation engine, for math, physics, and computer science papers, and it would work by learning the preferences of the user.

The metadata for each article includes an abstract, which should contain enough information to determine the relevance of the article. We decided that performing text classification using the words in the abstracts would give us the feature data needed to create a learning algorithm that would present interesting papers to the user and filter out those that fall outside the user's research interest. Furthermore it seemed most useful that the tool employ online learning, receiving a paper, extracting feature data and obtaining labels from the user, and updating the hypothesis one article at a time, so that the tool can present abstracts while getting progressively more accurate.

One of the important aspects of the algorithm is that it present some articles predicted to be uninteresting. Otherwise there is risk in mislabeling and potentially discarding many articles that are interesting to the user. Nevertheless the goal is to have an active filter on the articles lest the whole purpose of the program be defeated. Since there must be a balance between presenting only interesting articles and having a full understanding of the user's interests, we must choose what fraction of uninteresting articles are presented. In the following section we present a bound on the number of updates made by the Perceptron algorithm we employ, justifying why the algorithm may work towards an accurate hypothesis while receiving labels for only a fraction of the samples labeled uninteresting.

In Section 3, we describe the main components of the program in the python modules we created, and give a description on how to use the tool. We present our results in Section 4, followed by other observations and concluding remarks.

%it is important that not all papers that are predicted uninteresting be discarded by the engine. Without the true there is risk in

\section{Theory: Algorithm and Bounding Updates}

The main engine behind the tool is a modified Perceptron algorithm which employs an accept-reject method among the papers it labels uninteresting. The following pseudocode details the algorithm, where $x_t$ denotes feature vectors, $y_t$ denotes labels, $\widehat{y}_t$ are predictions, $w_t$ are hypotheses, and $k_t$ are i.i.d. random variables used for the accept-reject method. $k_t = 1$ with probability $p$ and $k_t = 0$ with probability $1-p$. As mentioned, if the algorithm only made updates when $\widehat{y}=1$ (i.e. interesting) then this would leave the number of mistakes made when $\widehat{y} = -1$ unchecked (hence, several lost interesting articles). As a compromise, we generate the i.i.d. variables $k_t$ for each observation when $\widehat{y} = -1$. Note that if $\widehat{y} = -1$ and $k=1$, the algorithm will ask for the correct label and make an update. Otherwise if $\widehat{y} = -1$ and $k=0$, the algorithm does not ask for the correct label and makes no update.\\

\noindent Modified Perceptron algorithm:
\begin{quotation}
%\noindent Modified Perceptron:\\
\noindent $w_1=0$\\
{\tt for $t=1$ to $T$}:\\
\indent receive($x_t$)\\
\indent $\widehat{y}_t=\mbox{sign}(w_t\cdot x_t)$\\
\indent {\tt if} $\widehat{y}_t =1$ {\tt or} ($\widehat{y}_t =-1$ {\tt and} $k_t = 1$):\\
\indent\indent receive($y_t$)\\
\indent\indent {\tt if $y_t\neq\widehat{y}_t$}:\\
\indent\indent\indent $w_{t+1} = w_t + y_t x_t$\\
\indent\indent {\tt else $w_{t+1} = w_t$}\\
\indent {\tt else if $\widehat{y}_t = -1$ and $k_t=0$}:\\
\indent\indent $w_{t+1} = w_t$\\
{\tt return $w_{T+1}$}

\end{quotation}

We learned in the course how it can be shown that the number of updates occuring in the original Perceptron algorithm can be bounded in terms of the separating margin $\rho$ and the size $r$ of the ball containing the feature data. We wanted to find a similar bound in our case where the number of updates is limited by how often we ask for the correct label when $\widehat y = 0$. Naturally we should aim for a probabilistic bound as opposed to a concrete bound since we incorporate the extra random variables $k_t$.

Let $I_1$ be the set of $t$ where a mistake was made and
$\widehat{y}=1$ (false positives).  Let $I_2$ be the set where a mistake was made with
$\widehat{y} = -1$ and $k = 1$ (false negatives where the algorithm updates), and $I_3$ the set where a mistake was
made with $\widehat{y} = -1$ and $k = 0$ (false negatives where the algorithm does {\it not} make an update). Finally, let $M_i = |I_i|$ for $i =
1,2,3$. We bound $M_1 + M_2$ in the same way we bound $M$ in the original Perceptron algorithm. Assume there exists a $\rho$ and $R$ satisfying the assumptions of Theorem 7.8 from our text, and a separating hyperplane determined by the vector $v$. Let $M' = M_1 + M_2$, then
\begin{align*}
M' \rho =&\ \frac{v}{\|v\|}\sum_{I_1 \cup I_2} y_tx_t \
 \leq   \Bigl\|\sum_{t \in I_1 \cup I_2} y_t x_t\Bigr\| \\
=& \ \Bigl\|\sum_{t \in I_1 \cup I_2} (w_{t+1} - w_t) \Bigr\| 
= \ \|w_{T+1} \| \\
=&\ \sqrt{\sum_{t \in I_1 \cup I_2} \|  w_{t+1}\|^2 - \|w_t\|^2 }\\
= &\ \sqrt{\sum_{t \in I_1 \cup I_2} \|  w_{t} + y_t x_t\|^2 - \|w_t\|^2 }\\
= &\ \sqrt{\sum_{t \in I_1 \cup I_2} 2y_t w_t\cdot x_t - \|x_t\|^2} \\
\leq &\ \sqrt{\sum_{t \in I_1 \cup I_2} \|x_t\|^2}
\leq \sqrt{M'r^2}.
\end{align*}
This implies that $M' \leq \frac{r^2}{\rho^2}$. Since $M = M' + M_3$, a bound on $M_3$ is needed. We revert to a weaker bound and consider the expectation of
$M_3$. Since 
\[M_3 = \sum_{t=1}^T 1_{\left\{\widehat{y}_t = -1 \cap y_t = 1\right\}}1_{k_t = 0}, \]
we get that
\begin{align*}
\E[M_3] = & \sum_{t=1}^T \E[1_{\left\{\widehat{y_t} = -1 \cap y_t = 1\right\}}]\E[1_{k_t = 0}] \\
 = &\ (1 - p)\sum_{t=1}^T \E[1_{\left\{\widehat{y_t} = -1 \cap y_t = 1\right\}}],
\end{align*}
where the last equality uses the independence of $k_t$ from the data. To control the sum of expectations in this expression we consider the expectation of $M_2$. Since
\[
M_2 =  \sum_{t=1}^T 1_{\left\{\hat{y_t} = -1 \cap y_t = 1\right\}}1_{k_t = 1},
\]
then
\begin{align*}
\E[M_2] =&\ \sum_{t=1}^T \E[1_{\left\{\hat{y_t} = -1 \cap y_t = 1\right\}}]\E[1_{k_t = 1}] \\
=&\ p\sum_{t=1}^T \E[1_{\left\{\hat{y_t} = -1 \cap y_t = 1\right\}}].
\end{align*}
Thus
\[
E[M_3] = \ \frac{1-p}{p}\ \E[M_2].
\]
We can apply the crude bound $\E[M_3] \leq \frac{1-p}{p}\ \E[M']$, since $M_2 \leq M'$. Thus the expected number of mistakes can be bounded by
\begin{align*}
\E[M]  =  &\ \E[M'] + \E[M_3] \\
\leq & \ \frac{r^2}{\rho^2} + \frac{1-p}{p}\frac{r^2}{\rho^2} \\
\leq  & \ \frac{1}{p}\frac{r^2}{\rho^2}
\end{align*}
Note that this is a weak bound.  Not only is it greater than the
original perceptron bound by a factor of $\frac{1}{p}$, it is also only a
bound in expectation.

\section{Implementation}

The tool is made from components in multiple Python files, split up according to their function in the program.
The metadata for each article is obtained using arXiv's Open Archives Initiative protocol which returns an XML file with a list of up to 1000 articles as nodes containing a variety of metadata, including abstracts, authors, submission dates, and fields and subfields, making the data quite accessible and fairly simple to sort.
The file {\tt ArxivData.py} contains the functions and class definitions used to collect, organize and process data. Included are a class of objects representing papers and a class definition for feature vectors.
%
%DESCRIBE BAG OF WORDS APPROACH FOR FEATURE VECTORS SOMEWHERE
%
The primary function is {\tt GetPapersOAI()} and its wrapper which prompts the user for a particular subject, any of arXiv's listed subfields for that subject, and a range of dates from which to select the papers.
Then it draws the XML file of data from arXiv's database, filters the articles using the XML entries according to the subfields, and prepares the stream of abstracts to present the user.
The user is then asked to label each article as interesting or not, one at a time, with updates made according to the algorithm employed.

%To use the program, run both scripts {\tt getTestData.py} to collect the papers to rate and then {\tt LabelData.py}

The files {\tt Perceptron.py} and {\tt OnlineSVM.py} house the classes that compose the different classifying algorithms we use.
The classes contain the methods that run the algorithm and the attributes such as weights and parameters that the algorithm uses and updates.
Within {\tt Perceptron.py} is the modified Perceptron algorithm we presented earlier, a Margin Perceptron classifier, and a Kernel Perceptron classifier supplied with both a polynomial and Gaussian kernel.
Similarly, the Online SVM algorithm is implemented in the file {\tt OnlineSVM.py}.
Each algorithm takes a learning rate $\eta$ as a parameter. The modified Perceptron requires no extra parameters, while the Gaussian Kernel Perceptron classifier requires a $\sigma$ specifying the variance in the kernel.
The Polynomial Kernel Perceptron takes two parameters $c$ and $d$ which are the constants used in constructing the polynomial kernel.
The Margin Perceptron requires a parameter $\rho$ designating the separating margin of the feature vectors, and the Online SVM algorithm takes a parameter $C$ for the penalty applied to the update (i.e. $w_{t+1} = w_t -\eta(w_t- Cy_tx_t)$).
Each algorithm class has an associated wrapper that instantiates a class, and we use each of these algorithms when testing the tool.

There are some additional Python files included that we used primarily for testing. For instance, {\tt getTestData.py} and {\tt LabelData.py} are scripts that call the functions that collect arXiv data and present the abstracts to acquire labels, respectively, which we place in a single script for the user. Additionally {\tt ClassifierTester.py} defines a class for the classification algorithms we include allowing them to be used in a uniform manner despite taking different parameters. This enables us to run tests over each algorithm, with multiple data sets from a few different users, using a set of several values for each parameter in a fairly organized manner in the script {\tt TestClassifiers.py}. Plots are also constructed in that script as well.

One of the most important aspects of this tool is the feature data itself, and what characteristics distinguish

\end{document}

%Recommendation engines have become an important feature of many services on the internet, and

% We envisioned that eventually after receiving enough data the tool can tailor the articles it presents to the preferences of the user and offer mostly articles that are ``interesting." The tool would sometimes offer articles it predicts that are uninteresting in order to get a full

\end{document}